---
title: "tfNeuralODE-Spirals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tfNeuralODE-Spirals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette runs through a small example of how to use tfNeuralODE to learn differential equations. We start by importing all of the correct packages, and instantiating a few constants.

```{r}
# initial checks for python
library(reticulate)
if(!py_available()){
  install_python()
}
library(tensorflow)
#checking for tensorflow and keras installation
if(!py_module_available("tensorflow")){
  install_tensorflow()
}

library(tfNeuralODE)
library(keras)
library(deSolve)

# constants
data_size = 1000
batch_time = 20 # this seems to works the best ...
niters = 200
batch_size = 16
```

Now we create the neural network that will define our system. We also instantiate a few more constants that define our system.

```{r}
# ODE Model with time input

OdeModel(keras$Model) %py_class% {
  initialize <- function() {
    super$initialize()
    self$block_1 <- layer_dense(units = 50, activation = 'tanh')
    self$block_2 <- layer_dense(units = 2, activation = 'linear')
  }

  call <- function(inputs) {
    x<- inputs ^ 3
    x <- self$block_1(x)
    self$block_2(x)
  }
}
model<- OdeModel()
# more constants, time vectors
tsteps <- seq(0, 25, by = 25/data_size)
true_y0 = c(2., 0.)
true_A = rbind(c(-0.1, 2.0), c(-2.0, -0.1))
```

Now we solve the ODE and plot our results.

```{r}
# solving a spiral ode
trueODEfunc<- function(du, u, p, t){
  true_A = rbind(c(-0.1, 2.0), c(-2.0, -0.1))
  du <- (u^3) %*% true_A
  return(list(du))
}

# solved ode output
prob_trueode <- lsode(func = trueODEfunc, y = true_y0, times = tsteps)

# plotting results
plot(prob_trueode[,2], prob_trueode[,3], type = "l")
```

Let's start looking at how we're going to train our model. We instantiate an optimizer and create a batching function.

```{r}
#optimizer
optimizer = tf$keras$optimizers$legacy$Adam(learning_rate = 1e-3)

# batching function
get_batch<- function(prob_trueode, tsteps){
  starts = sample(seq(1, data_size - batch_time), size = batch_size, replace = FALSE)
  batch_y0 <- as.matrix(prob_trueode[starts,])
  batch_yN <- as.matrix(prob_trueode[starts + batch_time,])
  batch_y0 <- tf$cast((batch_y0), dtype = tf$float32)
  batch_yN <- tf$cast((batch_yN), dtype = tf$float32)
  return(list(batch_y0, batch_yN))
}
```

Now we can train our neural ODE, using naive backpropagation. This process will output a plot every 100 iterations. This is not run by default, due to the length of time needed.
```{r, eval=FALSE}
# Training Neural ODE
for(i in 1:niters){
  #print(paste("Iteration", i, "out of", niters, "iterations."))
  inp = get_batch(prob_trueode[,2:3], tsteps)
  pred = forward(model, inputs = inp[[1]], tsteps = tsteps[1:batch_time])
  with(tf$GradientTape() %as% tape, {
    tape$watch(pred)
    loss = tf$reduce_mean(tf$abs(pred - inp[[2]]))
  })
  #print(paste("loss:", as.numeric(loss)))
  dLoss = tape$gradient(loss, pred)
  list_w = backward(model, tsteps[1:batch_time], pred, output_gradients = dLoss)
  optimizer$apply_gradients(zip_lists(list_w[[3]], model$trainable_variables))

  # graphing the Neural ODE
  if(i %% 200 == 0){
    pred_y = forward(model = model, inputs = tf$cast(t(as.matrix(true_y0)), dtype = tf$float32),
                     tsteps = tsteps, return_states = TRUE)
    pred_y_c<- k_concatenate(pred_y[[2]], 1)
    p_m<- as.matrix(pred_y_c)
    plot(p_m, main = paste("iteration", i), type = "l", col = "red")
    lines(prob_trueode[,2], prob_trueode[,3], col = "blue")
  }
}
```

